Study 2 Analysis

### 2024-12-18 nb/s1_survey.qmd
update - marginal effects

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc,els) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

>>>>

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc,els) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ "0.01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

### 2024-12-18 nb/s2_survey.qmd
update - marginal effects

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded,els) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

@@

s2_agg4 |> group_by(accuracy_level) |> summarize(n=n())
s2_agg1 |> group_by(accuracy_level) |> summarize(n=n())

# Calculate correlations
cor.test(s2_agg4$els, s2_agg4$mean_abs_error)
cor.test(s2_agg4$els, s2_agg4$mean_log_abs_error)
cor.test(s2_agg1$els, s2_agg1$log_abs_error)

@@

m_equal <- clm(accuracy_level ~ els, 
    data = s2_agg1, 
    link="probit")

@@

# Fit a mixed-effects model with accuracy level as the outcome
library(lme4)
mixed_model <- lmer(abs_error ~ els + (1 | id), data = s2_agg1)

@@

polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s2_agg1)
summary(polr_model_ref_interaction)

# Example for refClass = USD, calc = No Calculator
new_data <- expand.grid(els = seq(min(s2_agg1$els), max(s2_agg1$els), length.out = 100),
                        refClass = "USD",
                        calc = "No Calculator")

@@

# Reshape to long format for ggplot
predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
                                            cols = `Exact match`:`Over 15% error`,
                                            names_to = "accuracy_level",
                                            values_to = "probability")

@@

s2_agg <- s2_agg1 %>%
  mutate(els_quartile = ntile(els, 4))

@@

s2_els_log_error <- brm(
    log_abs_error ~ els + (1|id) + (1|state),
    data = s2_agg1,
    family = gaussian(),
    cores = 4,
    iter = 3000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error.rds')) 
)

@@

s2_els_log_error_full_add <- brm(
    log_abs_error ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    family = gaussian(),
    cores = 4,
    iter = 4000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error_full_add.rds')) 
)
summary(s2_els_log_error_full_add)

check_model(s2_els_log_error_full_add)

s2_els_acc_full_add <- brm(
    accuracy_level ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_error_full_add.rds')) 
)
summary(s2_els_acc_full_add)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -5.67      0.84    -7.31    -4.01 1.00     1039     1978
# Intercept[2]          -4.22      0.83    -5.84    -2.59 1.00     1032     2049
# Intercept[3]          -0.50      0.80    -2.04     1.08 1.00      964     1707
# els                   -6.38      0.95    -8.29    -4.59 1.00      972     1605
# refClassPercentage     0.55      0.58    -0.57     1.66 1.00      950     1652
# refClassUSD            1.07      0.62    -0.13     2.31 1.00      844     1563
# calcNoCalculator       3.18      0.67     1.90     4.55 1.01      872     1585
# roundedRounded        -0.65      0.18    -0.99    -0.31 1.00     5287     3387
# pct_goal15%           -0.43      0.17    -0.76    -0.08 1.00     6491     3227

@@

s2_els_acc_full_int1 <- brm(
    accuracy_level ~ (els*refClass)+(rounded*pct_goal)+calc + (1|id) + (1|state),
    data = s2_agg1,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_full_int1.rds')) 
)
summary(s2_els_acc_full_int1)
conditional_effects(s2_els_acc_full_int1)
# Intercept[1]                  -5.64      0.92    -7.48    -3.82 1.00     1119     2097
# Intercept[2]                  -4.19      0.91    -6.00    -2.36 1.00     1084     2079
# Intercept[3]                  -0.45      0.90    -2.23     1.31 1.00     1074     1995
# els                           -6.13      1.10    -8.30    -3.95 1.00     1140     2308
# refClassPercentage             0.97      0.96    -0.87     2.90 1.00     1139     1717
# refClassUSD                    1.08      1.02    -0.90     3.11 1.01     1066     1943
# roundedRounded                -0.88      0.48    -1.83     0.02 1.00      795     1583
# pct_goal15%                   -0.66      0.48    -1.60     0.24 1.01      809     1652
# calcNoCalculator               3.18      0.68     1.86     4.55 1.00      956     1621
# els:refClassPercentage        -0.66      1.32    -3.36     1.89 1.00     1270     1956
# els:refClassUSD               -0.10      1.41    -2.83     2.64 1.00     1650     2359
# roundedRounded:pct_goal15%     0.46      0.90    -1.22     2.20 1.00      737     1534

>>>>

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded,els) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match","0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ ".01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

@@

s2_agg4 |> group_by(accuracy_level) |> summarize(n=n())
s2_agg |> group_by(accuracy_level) |> summarize(n=n())

# Calculate correlations
cor.test(s2_agg4$els, s2_agg4$mean_abs_error)
cor.test(s2_agg4$els, s2_agg4$mean_log_abs_error)
cor.test(s2_agg$els, s2_agg$log_abs_error)

@@

m_equal <- clm(accuracy_level ~ els, 
    data = s2_agg, 
    link="probit")

@@

# Fit a mixed-effects model with accuracy level as the outcome
library(lme4)
mixed_model <- lmer(abs_error ~ els + (1 | id), data = s2_agg)

@@

polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s2_agg)
summary(polr_model_ref_interaction)

# Example for refClass = USD, calc = No Calculator
new_data <- expand.grid(els = seq(min(s2_agg$els), max(s2_agg$els), length.out = 100),
                        refClass = "USD",
                        calc = "No Calculator")

@@

# Reshape to long format for ggplot
predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
                                            cols = `Exact match`:`Over 5% error`,
                                            names_to = "accuracy_level",
                                            values_to = "probability")

@@

s2_agg <- s2_agg %>%
  mutate(els_quartile = ntile(els, 4))

@@

s2_els_log_error <- brm(
    log_abs_error ~ els + (1|id) + (1|state),
    data = s2_agg,
    family = gaussian(),
    cores = 4,
    iter = 3000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error.rds')) 
)

@@

s2_els_log_error_full_add <- brm(
    log_abs_error ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg,
    family = gaussian(),
    cores = 4,
    iter = 4000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error_full_add.rds')) 
)
summary(s2_els_log_error_full_add)

#check_model(s2_els_log_error_full_add)

s2_els_acc_full_add <- brm(
    accuracy_level ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_error_full_add.rds')) 
)
summary(s2_els_acc_full_add)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -5.67      0.84    -7.31    -4.01 1.00     1039     1978
# Intercept[2]          -4.22      0.83    -5.84    -2.59 1.00     1032     2049
# Intercept[3]          -0.50      0.80    -2.04     1.08 1.00      964     1707
# els                   -6.38      0.95    -8.29    -4.59 1.00      972     1605
# refClassPercentage     0.55      0.58    -0.57     1.66 1.00      950     1652
# refClassUSD            1.07      0.62    -0.13     2.31 1.00      844     1563
# calcNoCalculator       3.18      0.67     1.90     4.55 1.01      872     1585
# roundedRounded        -0.65      0.18    -0.99    -0.31 1.00     5287     3387
# pct_goal15%           -0.43      0.17    -0.76    -0.08 1.00     6491     3227

@@

s2_els_acc_full_int1 <- brm(
    accuracy_level ~ (els*refClass)+(rounded*pct_goal)+calc + (1|id) + (1|state),
    data = s2_agg,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_full_int1.rds')) 
)
summary(s2_els_acc_full_int1)
conditional_effects(s2_els_acc_full_int1)
# Intercept[1]                  -5.64      0.92    -7.48    -3.82 1.00     1119     2097
# Intercept[2]                  -4.19      0.91    -6.00    -2.36 1.00     1084     2079
# Intercept[3]                  -0.45      0.90    -2.23     1.31 1.00     1074     1995
# els                           -6.13      1.10    -8.30    -3.95 1.00     1140     2308
# refClassPercentage             0.97      0.96    -0.87     2.90 1.00     1139     1717
# refClassUSD                    1.08      1.02    -0.90     3.11 1.01     1066     1943
# roundedRounded                -0.88      0.48    -1.83     0.02 1.00      795     1583
# pct_goal15%                   -0.66      0.48    -1.60     0.24 1.01      809     1652
# calcNoCalculator               3.18      0.68     1.86     4.55 1.00      956     1621
# els:refClassPercentage        -0.66      1.32    -3.36     1.89 1.00     1270     1956
# els:refClassUSD               -0.10      1.41    -2.83     2.64 1.00     1650     2359
# roundedRounded:pct_goal15%     0.46      0.90    -1.22     2.20 1.00      737     1534
